Automatic Identification and Annotation of Tasks in Structured Programs

### 2018-02-19 docs/ACM/paper.tex
Section 3 finally concluded.

%\pedro{In the Kenjiro's paper: "This scheduling policy suffers in two situations: 
%(1) the parent is a long for loop creating a large number of children, 
%all these children cannot start until the parent finishes the task-generating loop (SparseLU); 
%(2) binary recursions that, rightly, spawn only the first recursive call as a task (FFT). 
%It imposes an upper bound for the ready task queue of a worker; a worker 
%is discouraged to create too many tasks ready on the queue"}.

To solve this problem, we conservatively limit the number of tasks
that can be created tied to the same Task Region. We do that by two means:
(1) limiting the creation of recursive tasks by depth; (2) limiting the creation of 
tasks in the whole program.

Consider the FIB

\pedro{Show FFT example}

\pedro{Explain that the setting is conservative and thus the parameter of control
(max recursion depth and total number of tasks) can be easily switched by the
user, or even completely disabled.}

@@

@@

>>>>

To solve this problem, we conservatively limit the number of tasks
that can be created tied to the same Task Region. We do that by two means:
(1) limiting the creation of recursive tasks by depth; (2) limiting the creation of 
tasks in the whole program. We do that by setting up static variables to count up
the depth level in recursive calls. When the depth reaches a certain level, which
can be defined by the user, task dispatch will be halted, avoiding creation of tasks that are
too small. This technique is illustrated in Figure \ref{fig:ex_cutoff}. 
This parameter's default value is very conservative and so can be easily switched off
or set by the user.

@@

As described in Challenge \ref{ch:privatize}, the need for privatization of variables
arises when tasks write on or read from variables that are shared among tasks but, instead,
should be private for each task. To distinguish variables that should be replicated among tasks
from those who shouldn't, we mapped this problem into a simple flow analysis, described
as follows.

\begin{definition}[Privatization]
\label{def:privatize}
A variable $X$ should be made private in a Task Region $T$ if the 
\emph{Definition-Use Chain} of $X$ crosses
the frontier of the task creation of $T$.
\end{definition}

This means that, if a variable $X$ is defined prior to the entry point of $T$, and then
it is used (read/write) inside $T$, then it should be made private. Figure \ref{fig:ex_privatize}
shows an example of privatization.

We only privatize values, not pointers. And for a matter of conservativeness, we set all the
variables shared by default, and then we run the privatization analysis afterwards to pin out
variables that should be made private.

@@

All the analyses and techniques described previously in this section are made upon
an \emph{intermediate representation} (IR) of the program. None of our work would
be made possible if we couldn't recover information from the IR back to Source Code.
{\Taskminer} needs not only to insert \emph{pragmas}
around task regions, but also recover array bounds information and preserve
soundness regarding scope. The bridge between IR and Source Code information
has 3 main obstacles to be crossed over:

\begin{compactitem}
\item \textsf{Source Code mapping}: The first problem is, given a set of instructions in the IR,
how to map them back to the source code file. {\Taskminer} compiles the source code
with debugging flags on, in order to extract debugging information that maps
instructions to lines of code. With these lines of code in hands, we then have to deal
with two more problems before annotation: scope breaking and array bounds information.

\item \textsf{Array bounds}: To recover array bounds information, we rely on strong 
alias analyses \pedro{References here?}. We link several kinds of instructions such as
loads, stores, array indexing and other pointer arithmetic instructions to 
trace back information to Source Code. We build a symbolic range analysis
for each pointer, map them to all loads and stores in the program and then come up with
lower and upper bounds for each. If we cannot properly define the range for all pointers 
inside a task region, \emph{we don't annotate the task.}

\item \textsf{Scope}: Once the task region is defined in terms of basic blocks in 
the \emph{Control Flow Graph} in the IR, a problem of scope breaking arises when
inserting \emph{pragmas} around code fragments in the Source Code. To work around
this problem, we build a \emph{Scope Tree} in LLVM's front-end (Clang) and before
we annotate the tasks, we check if the region mapped back to Source Code will break
any scope in the original program. The scope-breaking risk appears when many instructions
inside basic blocks map back to their definition (instead of their uses) in the source code,
which sometimes are outsite the region defined as a task. When that happens, we opt 
not to annotate the task, because we can't recover enough information to correctly annotate
the original code.

\end{compactitem}
